\section{Types of Collaboration}
\label{appendix:types-collaboration}

Table~\ref{tab:coding-types-collaboration} presents the systematic coding of collaboration types and characteristics across all 33 papers. This coding sheet addresses RQ1 by examining the various forms of collaborative AR implementation, interaction mechanisms, and application contexts in industrial settings.

\afterpage{
\clearpage
\begin{landscape}
\tiny
\begin{longtable}{@{}p{1.8cm}p{1.8cm}p{1.8cm}p{1.8cm}p{1.8cm}p{1.8cm}p{1.8cm}p{1.8cm}@{}}
\caption{Coding sheet for Types of Collaboration (RQ1)} \label{tab:coding-types-collaboration} \\
\toprule
\textbf{Paper} & \textbf{Application Type} & \textbf{Collaboration Type} & \textbf{Interaction Mechanism} & \textbf{Application Context} & \textbf{Task Complexity} & \textbf{Evaluation Metrics} & \textbf{Use Case Effectiveness} \\
\midrule
\endfirsthead

\caption[]{Coding sheet for Types of Collaboration (RQ1) (continued)} \\
\toprule
\textbf{Paper} & \textbf{Application Type} & \textbf{Collaboration Type} & \textbf{Interaction Mechanism} & \textbf{Application Context} & \textbf{Task Complexity} & \textbf{Evaluation Metrics} & \textbf{Use Case Effectiveness} \\
\midrule
\endhead

\midrule
\multicolumn{8}{r}{\textit{Continued on next page}} \\
\endfoot

\bottomrule
\endlastfoot

Aschenbrenner et al. (2018) & Knowledge Transfer \& Expert Support & Co-located multi-user with remote expert & First-person video perspective, 3D object manipulation & Industrial production planning & High—step-by-step guidance & Task duration, error rate, usability ratings & Demonstrated improvements in efficiency and accuracy \\
\midrule
Aschenbrenner et al. (2019) & Knowledge Transfer \& Expert Support & Single-user and collaborative repair tasks & Visual annotations, audio/text instructions & Industrial robot maintenance & High—complex repair steps & Task duration, error rate, task load & Projection-based SAR significantly improved performance \\
\midrule
Betti et al. (2018) & Collaborative Creative Workflows & Multi-user, co-located design with non-professionals & Hand gestures and voice commands & Architectural design and fabrication & Moderate—CAD manipulation & Ease of use, user engagement, error rate & Holographic AR enabled productive interaction \\
\midrule
Buyruk and Çağdaş (2022) & Collaborative Creative Workflows & Multi-user collaboration in shared MR environment & Hand gestures and voice commands & Parametric design and fabrication & High—real-time design updates & Real-time responsiveness, user adaptability & High user engagement and efficiency \\
\midrule
Chan et al. (2022) & Precision Task Execution & Human-robot collaborative manufacturing & Gesture, speech, and gaze inputs & CFRP fabrication & High—trajectory adjustments & Task load, system usability & Enhanced engagement, reduced physical demand \\
\midrule
Cogurcu et al. (2023) & Precision Task Execution & Human-robot collaboration & Safety zones via HoloLens & Pick-and-place tasks & Moderate—safety zone management & Trust, safety perception & Increased safety and task ease \\
\midrule
De Franco et al. (2019) & Precision Task Execution & Human-robot collaboration & Gesture, gaze, and voice & Industrial assembly and polishing & Moderate—sequential coordination & Task performance, cognitive load & Improved coordination with real-time feedback \\
\midrule
Fang et al. (2022) & Precision Task Execution & Multi-operator, distributed assembly & Visual guidance overlays & Collaborative assembly on shop floors & High—multi-user synchronization & Localization accuracy, task efficiency & Improved task alignment and guidance \\
\midrule
Gemito et al. (2023) & Precision Task Execution & Human-robot collaboration & Visual/audio cues, step-by-step guidance & Automotive manufacturing & High—complex assembly & Task accuracy, safety perception & 30\% reduction in assembly time \\
\midrule
Goepel and Crolla (2020) & Collaborative Creative Workflows & Multi-user co-located creative assembly & Visual overlays, flow pattern guides & Bamboo art installation & High—sculptural alignment & User satisfaction, engagement & High flexibility and engagement \\
\midrule
Gursch et al. (2018) & Collaborative Creative Workflows & Human-robot collaboration in logistics & Real-time task updates, conflict alerts & Intralogistics and scheduling & High—conflict resolution & Task efficiency, user focus & Improved logistics coordination \\
\midrule
Huy et al. (2017) & Knowledge Transfer \& Expert Support & Co-located HRI with awareness modes & Laser-projected cues, see-through AR & Industrial task coordination & Moderate—real-time monitoring & Task engagement, safety perception & Enhanced awareness and safety \\
\midrule
Liu et al. (2018) & Dynamic Problem-Solving & Multi-user AR platform & 3D models, gaze indicators & Boat customization and sales & Moderate—interactive adjustments & Customer satisfaction, engagement & Enhanced customer experience \\
\midrule
Liu et al. (2022) & Knowledge Transfer \& Expert Support & Human-machine collaboration & Real-time instructions, quality monitoring & Aerospace assembly & High—precision alignment & Assembly efficiency, error rate & Improved accuracy and efficiency \\
\midrule
Lotsaris et al. (2021) & Knowledge Transfer \& Expert Support & Human-robot collaboration & Task instructions, robot control & Automotive assembly & Moderate—robot navigation & Task completion efficiency & Enhanced task accuracy \\
\midrule
Lu et al. (2021) & Precision Task Execution & Multi-user AR-based visualization & 3D sand table, environment visualization & Industrial park planning & Moderate—data visualization & System performance, user satisfaction & High user engagement \\
\midrule
Martins et al. (2024) & Knowledge Transfer \& Expert Support & Multi-user, multi-robot system & Real-time DT display, trajectory viewing & Industrial multi-robot environments & High—task coordination & Situational awareness, scalability & Improves safety and coordination \\
\midrule
Michalos et al. (2018) & Precision Task Execution & Industrial HRC system & Real-time AR instructions, smartwatch control & Automotive assembly & High—safety zone management & Cycle time reduction, safety incidents & Reduced physical load on operators \\
\midrule
Mourtzis et al. (2021) & Collaborative Creative Workflows & Multi-user collaborative design & 3D hologram manipulation, gestures & Manufacturing design & High—part customization & Design cycle time, user engagement & Reduced design cycle time \\
\midrule
Otto et al. (2016) & Knowledge Transfer \& Expert Support & Multi-user dual reality system & Real-time assembly visualization & Production planning & High—ergonomic evaluation & Error rate reduction, cycle time & Enhanced collaborative decision-making \\
\midrule
Putri et al. (2024) & Knowledge Transfer \& Expert Support & Collaborative safety system & Real-time visual cues, training simulations & Manufacturing safety & Moderate—hazard detection & Incident rate, response time & Improved safety communication \\
\midrule
Reis et al. (2021) & Precision Task Execution & Multi-user VR/AR simulation & Digital twin navigation, 3D manipulation & Offshore oil and gas & High—critical task simulations & Safety incident reduction & Reduced human exposure to risks \\
\midrule
Restas et al. (2024) & Precision Task Execution & Multi-user collaborative AR/VR & Real-time model customization & Social manufacturing & Moderate to high—customization & User engagement, communication effectiveness & Supports efficient co-creation \\
\midrule
Rogeau and Rezaei Rad (2024) & Knowledge Transfer \& Expert Support & Multi-level human-robot system & Holographic instructions, trajectory visualization & Timber assembly & High—complex joint fitting & Assembly accuracy, error reduction & Enhances precision and safety \\
\midrule
Rubart et al. (2022) & Precision Task Execution & Multimodal AR platform & Shared dashboards, gesture/voice control & Smart manufacturing control room & Moderate to high—monitoring & Situational awareness, response time & Enhanced real-time data access \\
\midrule
Schmidt et al. (2022) & Precision Task Execution & AR interface for HRI & Real-time robot metrics visualization & Robotic cell management & High—real-time monitoring & Cycle time, energy consumption & Enhances task efficiency and safety \\
\midrule
Stacchio et al. (2023) & Collaborative Creative Workflows & Collaborative MR platform & Holographic document interaction & Assembly task management & Moderate—document access & Task efficiency, error reduction & Enhances on-site productivity \\
\midrule
Verlinden and Bekker (2017) & Knowledge Transfer \& Expert Support & Multi-stakeholder AR platform & Holographic overlays with metrics & Urban infrastructure design & High—environmental analysis & Material efficiency, stakeholder alignment & Enhances environmental impact understanding \\
\midrule
Vidal-Balea et al. (2020) & Collaborative Creative Workflows & Industrial AR for training & Shared AR visualization & Shipbuilding training & Moderate—assembly guidance & Communication delay, transmission latency & Enhances task comprehension \\
\midrule
Wang et al. (2022) - Cross-platform & Precision Task Execution & Cross-platform AR communication & Real-time AR annotations & Shipbuilding assembly & Moderate to high—multi-device sync & Annotation accuracy, transmission latency & Enhances assembly coordination \\
\midrule
Wang et al. (2022) - Multi-person & Dynamic Problem-Solving & Multi-person collaborative AR & Role-based user interactions & Assembly process evaluation & High—role-specific tasks & Usability, task accuracy & Enhances assembly evaluation \\

Yang et al. (2023) & Collaborative Creative Workflows & Multi-user, non-dyadic AR & Task-specific holograms, role-based instruction & Timber prefabrication & High—coordinated task sharing & System Usability Scale, task accuracy & Facilitates coordinated task management \\
\end{longtable}
\end{landscape}
\clearpage
} 